###I selected the Us States Production dataset from [http://vincentarelbundock.github.io/Rdatasets/]

#####Source
######Munnell, A. (1990) “Why has productivity growth declined? ######Productivity and public investment”, New England Economic Review, 3–22.
######Baltagi, B. H. and N. Pinnoi (1995) “Public capital stock and state productivity growth: further evidence”, Empirical Economics, 20, 351–359.
#####References
######Baltagi, Badi H. (2003) Econometric analysis of panel data, John Wiley and sons, http://www.wiley.com/legacy/wileychi/baltagi/.
#####fields :

state      |the state
-----------|-------------------------------------
year       |the year
pcap       | private capital stock
hwy        |highway and streets
water      | water and sewer facilities
util       |other public buildings and structures
pc         | public capital
gsp        | gross state products
emp       |labor input measured by the employment in           non–agricultural payrolls
unemp     | state unemployment rate
#####I took interest in the unemployment rates and began exploring that column of the data to see its characteristics, how it changed over the time frame of the data, and what its relationship to the other columns was. 

##### I am going to investigate the question of how unemployment rates changed over time from 1970 to 1985 using this dataset.
#####First I found the summary of the data

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(maps)
library(mapdata)
library(ggplot2)
library(stringr)
library(dplyr)
library(reshape2)
library(tidyr)
library(ggmap)
library(plotly)
csv_raw <- "https://raw.githubusercontent.com/aschwenker/MSDS_bridge/master/Produc.csv"
prod_df <- read.csv(csv_raw)
summary(prod_df)
```
#####Then I looked at the relationship between the employment in non-agricultural payrolls and unemployment fields using a scatter plot
```{r echo=FALSE}
plot(prod_df$unemp, prod_df$emp, main="employment, unemployment rate relationship", 
     xlab="unemployment rate ", ylab="employment in nonagricultural payrolls ",pch=19)
```

#####based on this scatter plot it would seem that employment in non-agricultural payrolls does not have a strong correlation with the unemployment rate. There are many factors that contribute to unemployment rates, and those factors vary greatly over 15 years, so I was not surprised to see very little relationship here. I next looked to see how the unemployment rate changed over time by state

```{r echo=FALSE}
ggplot(data=prod_df, aes(x=year, y=unemp,color=state))+geom_line()+ theme(panel.background = element_rect(fill = "lightblue",
                                                                                                          colour = "lightblue",
                                                                                                          size = 0.5, linetype = "solid"),
                                                                          axis.text.x = element_text(angle = 90, hjust = 1))
```

##### Then I wanted to look at the mean unemployment rate by year
```{r echo=FALSE}
prod_df$year<-as.character(prod_df$year)
ggplot(prod_df) +
  geom_boxplot(aes(y = unemp, x = year))
```

##### Next I wanted to see the distribution of unemployment rates across the states in 1970

```{r echo=FALSE}
year_subset<-c(1970)
prod_df_subset <- subset(prod_df,(prod_df$year %in% year_subset) ,select=c(state,year, unemp))
prod_df_subset$year<-as.character(prod_df_subset$year)
hist(prod_df_subset$unemp,col="lightblue",main="Frequency of unemployment rates in 1970 in the United States",labels=TRUE,xlab="1970 unemployment rates")

```

##### Then I wanted to look at the change in unemployment rates from 1970 to 1985 using a density map, so I subset the years and restructure the data so that I could find the change between years. I then created a column that showed the difference in unemployment rates between 1970 and 1985. 

```{r echo=FALSE}
year_subset<-c(1970,1985)
prod_df_subset <- subset(prod_df,(prod_df$year %in% year_subset) ,select=c(state,year, unemp))
prod_df_subset$year<-as.character(prod_df_subset$year)
head(prod_df_subset)
names(prod_df_subset)[1]<- paste("region")
restruct_prod_list<-prod_df_subset %>% spread("year","unemp")
restruct_prod_df<-lapply(restruct_prod_list,function(x) if(is.factor(x)) as.character(x) else x)

restruct_prod_df<- as.data.frame(restruct_prod_df)
restruct_prod_df$unemp_change <- restruct_prod_df$X1985 - restruct_prod_df$X1970
head(restruct_prod_df)
```


#####I also needed to change all values to uppercase and change the spaces to _ in order to make the data from both match to merge. I then merged it with the states dataset from map_data function. Unfortunately the states are displayed as fractured and I was unable to fix this before submitting. My attempt is below.
```{r echo=FALSE}

states<-map_data("state")
unique(states$region)
states$region<-sub(" ","_",states$region)
states<-(lapply(states, function(v) {
  if (is.character(v)) return(toupper(v))
  else return(v)
}))
unique(states$region)
states<- as.data.frame(states)
prod_map<-merge(states,restruct_prod_df)
state_base <- ggplot(data = states, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")
head(prod_map)
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
)

final_map <- state_base + 
  geom_polygon(data = prod_map, aes(fill = unemp_change), color = "white") +
  geom_polygon(color = "black", fill = NA) +
  theme_bw()+ditch_the_axes

final_map 
```

My final conclusion is that unemployment rates varied greatly over the 15 year time span, and were not directly correlated with the labor input measured by the employment in non–agricultural payrolls.