## I selected the nkill.bCountryYr dataset from [http://vincentarelbundock.github.io/Rdatasets/]
## this source database is summarized as follows:
>Global Terrorism Database yearly summaries
Description
The Global Terrorism Database (GTD) "is a database of incidents of terrorism from 1970 onward". Through 2015, this database contains information on 141,966 incidents.

## The csv I used includes:
>
year
integer year, 1970:2014.
methodology
an ordered factor giving the methodology / organization responsible for the data collection for most of the given year. The Pinkerton Global Intelligence Service (PGIS) managed data collection from 1970-01-01 to 1997-12-31. The Center for Terrorism and Intelligence Studies (CETIS) managed the project from 1998-01-01 to 2008-03-31. The Institute for the Study of Violent Groups (ISVG) carried the project from 2008-04-01 to 2011-10-31. The National Consortium for the Study of Terrorism and Responses to Terrorism (START) has managed data collection since 2011-11-01. For this variable, partial years are ignored, so methodology = CEDIS for 1998:2007, ISVG for 2008:2011, and START for 2012:2014.
nkill
number of confirmed fatalities for incidents in the given year, including attackers = sum(nkill, na.rm=TRUE) in the GTD incident data.
NOTE: nkill in the GTD incident data includes both perpetrators and victims when both are available. It includes one when only one is available and is NA when neither is available. However, in most cases, we might expect that the more spectacular and lethal incidents would likely be more accurately reported. To the exent that this is true, it means that when numbers are missing, they are usually zero or small. This further suggests that the summary numbers recorded here probably represent a slight but not substantive undercount.




```{r echo=FALSE}




library(ggplot2)
library(reshape2)
library(RCurl)
df <- read.csv(text=getURL("https://raw.githubusercontent.com/aschwenker/MSDS_bridge/master/nkill.byCountryYr.csv"))

getwd()
compare_values<-function(x,y){
  if (x>y){
    return(paste(deparse(substitute(x))," is larger than ",deparse(substitute(y))))
  }else if(y>x){
    return(paste(deparse(substitute(y))," is larger than ",deparse(substitute(x))))
  }else if(x==y){
    return(" both values are the same")
  }
  }
# if integer, make numeric -- if factor, make character
df[] <- lapply(df, function(x) if(is.integer(x)) as.numeric(x) else x)
df[]<-lapply(df,function(x) if(is.factor(x)) as.character(x) else x)
```
### Below is the Summary for the full dataset
``` {r echo=FALSE}
summary(df)
```
##Below is the mean and median for the full dataset for 2010 and 2015
I selected these years as I saw from the summary greatest means were seen in 2013 through 2015 and I wanted to compare a 5 year span so I selected 2010 and 2015. Probably with a better knowledge of the topic I could have made different inferences about this dataset.
```{r echo=FALSE }
mean_2010 <- mean(df$X2010,na.rm=TRUE)
print(paste("The mean for 2010 was",mean_2010))
median_2010 <- median(df$X2010,na.rm=TRUE)
print(paste("The median for 2010 was",median_2010))
mean_2015 <- mean(df$X2015,na.rm=TRUE)
print(paste("The mean for 2015 was",mean_2015))
median_2015 <- median(df$X2015,na.rm=TRUE)
print(paste("The median for 2010 was",median_2010))

names(df)[1] <- paste("Country")
```
##Below I subsetted the full dataset for just 2010 and 2015 in order to compare them using a boxplot. In order to run a box plot I needed to restructure the data. you can see from the two sample below how it was originally structured and then how it is structure after using the melt function from the reshape2 library I also changed the column headings to Country, Year, Killings
```{r echo=FALSE}

df_mean_compare <- subset(df,select=c(Country,X2010, X2015))
head(df_mean_compare,n=10)
df_plot_full<-melt(df_mean_compare,value.name = "killings")
names(df_plot_full)[1] <- paste("Country")
names(df_plot_full)[2] <-paste("Year")
names(df_plot_full)[3] <- paste("Killings")

df_plot_full[]<-lapply(df_plot_full,function(x) if(is.factor(x)) as.character(x) else x)
df_plot_full$Year<-gsub( "X", "",df_plot_full$Year)
head(df_plot_full, n=10)
ggplot(df_plot_full) +
  geom_boxplot(aes(y = Killings, x = Year)) + coord_flip()
```
##Then I sorted the original dataset in descending order based on 2015 and selected the top 10 rows
```{r echo=FALSE}
df_sort <-df[order(-df$X2015),][1:10,]
df_sort
df_sort[]<-lapply(df_sort,function(x) if(is.factor(x)) as.character(x) else x)

#used [[]] below. with only single did not create vector, made a sublist instead according to
#https://stackoverflow.com/questions/7070173/r-friendly-way-to-convert-r-data-frame-column-to-a-vector
countries_subset<-df_sort$Country

```
##Next I subsetted the original dataset for all the countries that were listed when sorted in descending order on the year 2015, and selected only those countries for 2010 and 2015 and printed summary, the mean, and the median for each year of the subset
```{r echo=FALSE}
countries_subset
dfsub <- subset(df,(df$Country %in% countries_subset) ,select=c(Country,X2010, X2015))
summary(dfsub)
subset_mean_2010 <- mean(dfsub$X2010,na.rm=TRUE)
print(paste("The subset mean for 2010 was",subset_mean_2010))
subset_median_2010 <- median(dfsub$X2010,na.rm=TRUE)
print(paste("The subset median for 2010 was",subset_median_2010))
subset_mean_2015 <- mean(dfsub$X2015,na.rm=TRUE)
print(paste("The subset mean for 2015 was",subset_mean_2015))
subset_median_2015 <- median(dfsub$X2015,na.rm=TRUE)
print(paste("The subset median for 2015 was",subset_median_2015))
```
###Then I compared the mean for each between the full and subsetted dataset
```{r echo=FALSE}
print("comparing means")
compare_values(mean_2010,subset_mean_2010)
compare_values(mean_2015,subset_mean_2015)
print("comparing medians")
compare_values(median_2010,subset_median_2010)
compare_values(median_2015,subset_median_2015)
```

###Next I switched out the value for Afghanistan to be AF and Cameroon to be CR
```{r}
"Afghanistan" %in% dfsub$Country
dfsub$Country[dfsub$Country=='Afghanistan']<-'AF'
"Afghanistan" %in% dfsub$Country
dfsub$Country[dfsub$Country=='Cameroon']<-'CR'
head(dfsub, n=10)
```
### Then I created a box plot for the subset
```{r echo=FALSE}

df2<-melt(dfsub,value.name = "killings")

names(df2)[1] <- paste("Country")
names(df2)[2] <-paste("Year")
names(df2)[3] <- paste("Killings")

df2[]<-lapply(df2,function(x) if(is.factor(x)) as.character(x) else x)
df2$Year<-gsub( "X", "",df2$Year)

ggplot(df2) +
  geom_boxplot(aes(y = Killings, x = Year)) + coord_flip()

```

